Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-15 abstract/abstract.tex
Updates from Overleaf

Music, an integral part of human culture, offers a diverse yet complex field of study due to its intricate styles, mutable ground truth, and subjective nature. This research, conducted in collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), employs self-supervised contrastive learning of musical representation to uncover the fundamental structure of Western tonal music.

@@

Results XXXXXX

>>>>

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time. This idea parallels the nature of the symbolic domain, which maintains its essence despite being interpreted through diverse performances using various instruments and styles.

This research, conducted in collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), employs self-supervised contrastive learning of musical representation to uncover the fundamental structure of Western tonal music.

@@

Results XXXXXX TBC

### 2023-05-15 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

\section{Music Structure Analysis as downstream task}

@@

\input{figures/neural networks/r128 sphere}

@@

\begin{enumerate}
\item \textbf{Load the audio data}: The audio data is loaded into memory as a one-dimensional array, denoted as $\text{input\_audio}$, of length $N$.

>>>>

\section{Music Structure Analysis (MSA) as downstream task}

@@

@@

\begin{enumerate}
\item \textbf{Load the audio data}: The audio data is loaded into memory as a one-dimensional array of length $N$.

### 2023-05-15 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to specific attributes such as sound quality, tempo, key, orchestration, and signal noise that evolve and unfolds through time. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations. The distinct tonal quality of a waveform, heavily influenced by the specific timbre of the instrument or production technology, adds complexity to the transcription of its sonic properties into accurate sheet music notation.

@@

For example, despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint so to say, derived from the propulsion of their melodic and harmonic contours through non-diatonic major thirds.

>>>>

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time such as melody, harmony, rhythm, form, motif, counterpoint, etc. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations.

@@

For example, despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint, so to say, derived from the propulsion of their melodic and harmonic contours through non-diatonic major thirds.

### 2023-05-15 introduction/Subsections/Motivation.tex
Updates from Overleaf

As a music enthusiast, professional musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them. This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, being my main interest in the audio and symbolic domain, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

On top of that, I argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond DSP.

\subsubsection{Academic motivation}

The emergence of self-supervised models that learn embedding spaces for retrieving musical content from audio signals has unveiled new opportunities in academic research. By capturing high-level semantic information from raw audio data in an unsupervised manner, these models can potentially revolutionize various aspects of the music industry. Their applications encompass music information retrieval, sharing learned latent representations for MIR \cite{HamelTransferSimilarity}, and fostering interdisciplinary research and innovation in musicology, computer science, and artificial intelligence.

\subsubsection{Industrial and corporate motivation}

In the industrial and corporate realm, the embedding spaces generated by these models can be utilized to create personalized and intelligent music recommendation systems \cite{Chen2020LearningRecommendation}\cite{epidemic}, facilitate effective enforcement of intellectual property rights, and contribute to the development of more sophisticated audio editing-production tools and technical products \cite{WonEmotionStories}. These advancements ultimately enhance our interaction with music in the digital age, promoting innovation and collaboration across the music industry.

@@

@@

Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception.

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within an existing tonal paradigm, such as Heinrich Schenker and his school of thought \cite{Komar1959SchenkersStructure}, particularly by challenging it, exemplified by Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell} or Ernst Levy \cite{LevyAHarmony}. These approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding tonal and atonal paradigms.

>>>>

As a music enthusiast, professional musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them.

This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, being my main interest in the audio and symbolic domains, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

\subsubsection{Academic and industrial motivation}

@@

The emergence of self-supervised models that learn embedding spaces for retrieving musical content from audio signals has unveiled new opportunities in academic research. By capturing high-level semantic information from raw audio data in an unsupervised manner, these models can potentially revolutionize various aspects of the music industry.

In the industrial and corporate realm, the embedding spaces generated by these models can be utilized to create all sorts of valuable products and user experiences for content creators. These advancements ultimately enhance our interaction with music in the digital age, promoting innovation and collaboration across the music industry.

@@

Purely mathematical approaches may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception. We argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond digital signal processing.

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within an existing tonal paradigm, such as Heinrich Schenker and his school of thought \cite{Komar1959SchenkersStructure}, and particularly by challenging it, exemplified by Arnold Schoenberg \cite{Samson1974SchoenbergsMusic}, George Russell \cite{LydianRussell} or Ernst Levy \cite{LevyAHarmony}. These approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding tonal and atonal paradigms.

### 2023-05-15 introduction/Subsections/Objectives.tex
Updates from Overleaf

This study aims to explore the core tonal structure of Western music by constructing and analyzing an embedding space applicable to subsequent Music Information Retrieval (MIR) tasks.

By employing a blend of MIR computational techniques and a musicological approach, it intends to examine a wide spectrum of musical works across diverse styles and genres. Crucial to this is the understanding of aural skills – perceptual and cognitive abilities essential for interpreting musical works through listening.

The research focuses on the learning of a single embedding, $e(M)$, for a music piece $M$ which encapsulates the underlying high-level content such as melody contour, harmony inter-relationships, and other holistic elements. The structuring of the embedding space will prioritize the clustering of similar content and sonic quality features while maintaining a clear distinction between them.

The goal is to create an embedding $e(M)$ that captures high-level musical content and remains agnostic to production features.

>>>>

This study explores music's core tonal structure by constructing and analyzing an embedding space applicable to subsequent MIR tasks.

By blending MIR computational techniques and a musicological approach, we intend to examine a broad spectrum of musical works across diverse styles and genres to mimic human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities.

The research focuses on learning a single embedding, $e(M)$, for a music piece $M$ which encapsulates the underlying high-level content such as melody contour, harmony inter-relationships, and other holistic elements. The structuring of the embedding space will prioritize the clustering of high-level musical content while maintaining agnostic to sonic quality features.

### 2023-05-15 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\input{figures/neural networks/pipeline}

@@

\begin{figure}[ht]
\centering
\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}
\caption{\small{Representation of the Average Pooling operation. Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the 2D input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.}}
\end{figure}

@@

\subsubsection{Positive sample generation}
Given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top.

@@

\item \textbf{Frequency-domain effects:} Frequency domain manipulation alters the frequency content or pitch of the signal.
\begin{itemize}
    \item Pitch shift: Changes the signal's pitch by $\Delta p \in [-1200, 1200]$ cents.
\end{itemize}

@@

\begin{enumerate}
\item Calculate the minimum and maximum chunk lengths in samples:
2\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}
\item Generate random chunk lengths $l_1, l_2, \ldots, l_{n-1}$ from the uniform distribution on the interval $[l_{min}, l_{max}]$. Calculate the final chunk length as:
\begin{equation}
l_n = L_A - \sum_{i=1}^{n-1} l_i
\end{equation}
where $L_A$ is the length of the anchor signal in samples.
\item Split the anchor signal $A$ into chunks $C_1, C_2, \ldots, C_n$ according to the calculated chunk lengths.
\item Shuffle the chunks randomly to get the permuted chunks $C_{\sigma(1)}, C_{\sigma(2)}, \ldots, C_{\sigma(n)}$, where $\sigma$ is a random permutation of indices from $1$ to $n$.
\item Concatenate the shuffled chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

Ensure that the length of the anchor and negative signals are the same, i.e.,

\begin{equation}
L_A = L_N,
\end{equation}

where $L_N$ is the length of the negative signal in samples.

@@

Schroff, F., Kalenichenko, D. and Philbin, J. at Google initially proposed and used triplet loss to learn face recognition of the same person at different poses and angles. \cite{Schroff2015FaceNet:Clustering}

A triplet loss function with a variable margin can be written as follows:

@@

In this equation:

>>>>

@@

\begin{figure}[ht]
\centering
\begin{equation}
y_i = \frac{1}{N} \sum_{j=1}^{N} x_{ij}
\end{equation}
\caption[Average Pooling operation]{\small{Representation of the Average Pooling operation. Here, $y_i$ is the output tensor after the average pooling process. Each element $x_{ij}$ is part of the 2D input tensor. $N$ is the count of elements in the second dimension of the input tensor, while $i$ and $j$ are used to iterate over the first and second dimensions of the tensors, respectively.}}
\end{figure}

@@

\subsubsection{Positive sample generation}
Given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of the chain.

@@

\item \textbf{Frequency-domain effects:} Frequency domain manipulation alters the frequency content or pitch of the signal.
\begin{itemize}
    \item Pitch shift: Changes the signal's pitch by $\Delta p \in [-1200, 1200]$ cents. Two-octave range, one octave higher to one octave lower.
\end{itemize}

@@

\begin{enumerate}
\item Calculate the minimum and maximum audio chunk lengths in samples:
2\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}
\item Generate random audio chunk lengths $l_1, l_2, \ldots, l_{n-1}$ from the uniform distribution on the interval $[l_{min}, l_{max}]$. Calculate the final audio chunk length as:
\begin{equation}
l_n = L_A - \sum_{i=1}^{n-1} l_i
\end{equation}
where $L_A$ is the length of the anchor signal in samples.
\item Split the anchor signal $A$ into audio chunks $C_1, C_2, \ldots, C_n$ according to the calculated audio chunk lengths.
\item Shuffle the audio chunks randomly to get the permuted slices $C_{\sigma(1)}, C_{\sigma(2)}, \ldots, C_{\sigma(n)}$, where $\sigma$ is a random permutation of indices from $1$ to $n$.
\item Concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

Schroff, F., Kalenichenko, D., and Philbin, J. at Google initially proposed and used triplet loss to learn face recognition of the same person at different poses and angles. \cite{Schroff2015FaceNet:Clustering}

@@

In this equation, triplet loss function with a variable margin:
